Stability and Chaos in Multi Agent Reinforcement Learning

### 2020-08-26 Documents/ChaoticDynamics/AAAI Submission/main.tex
Generalised Code to p-players
Included NR Method

@@

In this section we will establish some of the preliminaries which are
required to follow the subsequent sections. The first is 'Dynamical
Behaviours' \fb{maybe italics?} which considers the evolution of agent strategies as they
learn on an iterated game. This study aims to classify, based on the
agent parameters and the payoff matrices, which of these behaviours
will be seen. The second is 'Dynamics of Q-Learning'. These are a set
of equations which model the aforementioned strategy evolution. It is
on these dynamics that we will perform our analysis.

@@

    It has long been established \fb{add refs?} that, upon lifting the strong
    assumptions made by traditional game theory (such as the
    rationality of agents and complete information), player
    strategies can result in much more complex behaviour than
    convergence to a Nash Equilibrium (NE). In fact, this is even true
    on what would commonly be regarded as 'simple' games such as
    tic-tac-toe and prisoner's dilemma \cite{Galla2011,
      Sato2002}. These behaviours include: convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    shown in Figure \ref{fig::DynamicalBehaviours}. Of these, the most
    preferable is, of course, convergence to a unique equilibrium,
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}) and so MARL techniques should
    avoid this. It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders et al. refer to these as the
    memory and intensity of choice parameters respectively \fb{what are $\alpha$ and $\tau$ exactly?}. Agent 1 takes action $i$ with probability
    $x_i$ while Agent 2 takes action $j$ with probability $y_j$. If these actions are taken, the agents
    receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
    predict the expected behaviour of Q-Learning agents, as shown in Figure 
    \ref{fig::TuylsExperiments} \fb{discuss limitations as well}.

@@

\section{Stability and Chaos in Multi-Agent Reinforcement Learning}

@@

To verify our theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The purpose of this experiment was to determine a heuristic estimate for the Lyapunov exponent, which determines whether initial conditions which start close to one another will converge or diverge from each other. The former indicates a stable system (in the sense of Lyapunov) whilst the latter signals instability \fb{add refs to all of this?}.

>>>>

In this study, we will be considering the question: under what parameter selections are is a
multi-agent system, which is trained on an iterated game using Q-Learning, likely to converge to an
equilibrium as opposed to complex behaviours? In investigating this question, we will make the
following assumptions:

\begin{enumerate}
    \item There are a finite set of agent, though the length $p$ of the set is arbitrary. 
    \item The agents have a large, but discrete, strategy space: during the theoretical study, we
    will make the assumption that the number of actions $N \rightarrow \infty$.
    \item The agents are homogeneous. This requires that all agents have the same parameters and are
    trained using the same algorithm. Since this is typically the case in reinforcement learning
    studies, it is not an unreasonable assumption, but does present an interesting avenue for future
    work.
    \item The agents are trained on stateless, normal form games (i.e. the environment is static).
\end{enumerate}

The parameters that we will consider are:

\begin{enumerate}
    \item $\alpha \in [0, 1]$: the \em{step length}. Low values of $\alpha$ denote smaller
    updates. Heuristically we can consider this to be the memory of the agent: lower $\alpha$
    denotes longer memory.
    \item $\tau in [0, \infty)$: the \em{intensity of choice}, as termed by Sanders et al. 
    \cite{Sanders2018}. This is sometimes written as $\beta$ in the literature \cite{Sutton2018}.
    $\tau = 0$ results in all actions being selected with equal probability, regardless of their
    Q-value, whilst $\tau \rightarrow \infty$ results in the action with the highest Q-value chosen
    at every step. This may be considered the \em{exploration-exploitation parameter}. For greater
    details into these parameters, the interested reader should consult \cite{Sutton2018}.
    \item $\Gamma \in [-1, p-1]$: the \em{payoff correlation}. This term controls the choice of
    covariance matrix from which random payoff matrices are drawn. $\Gamma = -1$ indicates a zero
    sum game, in which the sum of payoffs for a given action across all agents is 0, resulting in a
    purely competitive game. $\Gamma = p-1$ gives a purely co=operative game in which all agents
    share the same payoffs. The manner in which a game is generated from the choice of $\Gamma$ is
    described in Section \ref{sec::Theory} and follows the same procedure as outlined in \cite{Sanders2018}.
\end{enumerate}

@@

In this section we will establish some of the preliminaries which are
required to follow the subsequent sections. The first is \em{Dynamical Behaviours} \fb{maybe italics?} 
which considers the evolution of agent strategies as they
learn on an iterated game. This study aims to classify, based on the
agent parameters and the payoff matrices, which of these behaviours
will be seen. The second is \em{Dynamics of Q-Learning}. These are a set
of equations which model the aforementioned strategy evolution. It is
on these dynamics that we will perform our analysis.

@@

    It has long been established that, upon lifting the strong
    assumptions made by traditional game theory (such as the
    rationality of agents and complete information), player
    strategies can result in much more complex behaviour than
    convergence to a Nash Equilibrium (NE). In fact, this is even true
    on what would commonly be regarded as 'simple' games such as
    tic-tac-toe and prisoner's dilemma \cite{Galla2011,
      Sato2002}. These behaviours include: convergence to a unique
    equilibrium (though not always to an NE), convergence to one of
    multiple equilibria, limit cycles and chaos. These behaviours are
    shown in Figure \ref{fig::DynamicalBehaviours}. Of these, the most
    preferable is, of course, convergence to a unique equilibrium,
    although it is still possible to study systems with multiple
    equilibria or limit cycles \cite{Strogatz2000}. However, it would
    be difficult to control systems whose dynamics are governed by
    chaos (though research into controlling chaos is ongoing and rife
    with opportunity \cite{Fradkov2009}) and so MARL techniques should
    avoid this. It would, therefore, be a useful endeavour to
    determine the conditions under which these sorts of behaviours
    arise.

@@

    Here, $\alpha$ and $\tau$ are the parameters of the agent; Sanders et al. refer to these as the
    memory and intensity of choice parameters respectively \fb{what are $\alpha$ and $\tau$ exactly?}. Agent 1 takes action $i$ with probability
    $x_i$ while Agent 2 takes action $j$ with probability $y_j$. If these actions are taken, the agents
    receive payoff $a_{ij}$ and $b_{ji}$ respectively. With these equations, it is possible to
    predict the expected behaviour of Q-Learning agents, as shown in Figure 
    \ref{fig::TuylsExperiments}. Upon examination of the derivation in \cite{Tuyls2006AnGames}, it
    is clear that the dynamics generated give the expected behaviour and, therefore , cannot account
    for transients. It is also interesting to note that the dynamics (\ref{eqn::EOM}) do not include
    the parameter $\gamma$, which moderates the Q-update. The implication is that this has no effect
    on the expected behaviour of the system. \textbf{Shall I include some tests to confirm this?
    It does turn out to be mostly true, the value of $\gamma$ doesn't affect the overall behaviour,
    but rather results in smaller or larger steps. What do you think?}.
    as well}.

@@

\section{Determination of Phase Transition Line} \label{sec::Theory}

@@

To verify our theoretical results, and to examine the underlying structure of stability and chaos in
Multi-Agent Q Learning, we perform a series of numerical experiments by varying the parameters $\Gamma$ and
$\alpha$ whilst keeping $\tau$ fixed. The purpose of this experiment was to determine a heuristic estimate for the Lyapunov exponent, which determines whether initial conditions which start close to one another will converge or diverge from each other. The former indicates a stable system (in the sense of Lyapunov) whilst the latter signals instability \cite{Strogatz2000}.
